---
title: "W. pollock maturity stage - multinomial - spectral orange peak height"
author: "Morgan Arrington"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(glmnet)
library(tidyr)
library(ggpubr)
library(reshape2)
library(caret)

# TIDY DATA
##Load in data

dat <- read.csv("C:/Users/marri/OneDrive/Documents/AFSC A&G Contract/Raman Spectroscopy/raman-spectroscopy/all_peakarea_data.csv")

#Add GSI
#gonad weight (g) / fish weight (g)) *100
dat$GSI <- dat$gonad_weight/dat$organism_weight*100

#Make HistoMaturity levels
dat$HistoMaturity <- factor(dat$HistoMaturity , levels = c("IMM","ND","DEV","VIT","PSWN","SWN","PSNT","SPN","SNT","OSNT"))

#Convert amplitude to height 1/sigma*sqrt(2pi)
dat <- dat%>%
  mutate(peak_1_height = peak_1_amplitude/(peak_1_sigma*sqrt(2*pi)), peak_2_height = peak_2_amplitude/(peak_2_sigma*sqrt(2*pi)), peak_3_height = peak_3_amplitude/(peak_3_sigma*sqrt(2*pi)), peak_4_height = peak_4_amplitude/(peak_4_sigma*sqrt(2*pi)), peak_5_height = peak_5_amplitude/(peak_5_sigma*sqrt(2*pi)), peak_6_height = peak_6_amplitude/(peak_6_sigma*sqrt(2*pi)), peak_7_height = peak_7_amplitude/(peak_7_sigma*sqrt(2*pi)), peak_8_height = peak_8_amplitude/(peak_8_sigma*sqrt(2*pi)), peak_9_height = peak_9_amplitude/(peak_9_sigma*sqrt(2*pi)), peak_10_height = peak_10_amplitude/(peak_10_sigma*sqrt(2*pi)),
peak_11_height = peak_11_amplitude/(peak_11_sigma*sqrt(2*pi)),
peak_12_height = peak_12_amplitude/(peak_12_sigma*sqrt(2*pi)),
peak_13_height = peak_13_amplitude/(peak_13_sigma*sqrt(2*pi)))

```

### Visualize peak area and peak height by maturity stage
 
```{r, message = F, echo = F, fig.width= 10}
#reshape and prepare data
dat_long_area <- dat %>%
  tidyr::pivot_longer(dat, cols=c(seq(53, 101, by=4)),names_to = "peak_num", values_to = "peak_area")

dat_long_area$peak_num <- factor(dat_long_area$peak_num, levels = c("peak_1_area", "peak_2_area", "peak_3_area", "peak_4_area", "peak_5_area", "peak_6_area", "peak_7_area", "peak_8_area", "peak_9_area", "peak_10_area", "peak_11_area", "peak_12_area", "peak_13_area"))

dat_long_height <- dat %>%
  tidyr::pivot_longer(dat, cols=c(106:118),names_to = "peak_num", values_to = "peak_height")

dat_long_height$peak_num <- factor(dat_long_height$peak_num, levels = c("peak_1_height", "peak_2_height", "peak_3_height", "peak_4_height", "peak_5_height", "peak_6_height", "peak_7_height", "peak_8_height", "peak_9_height", "peak_10_height", "peak_11_height", "peak_12_height", "peak_13_height"))

#facet_wrap by peak, maturity stage along y axis, peak area along x
oo_cols <- length(unique(dat_long_area$HistoMaturity))
oo_colors <- colorRampPalette(c("#2b9348","#B9E769","#EFEA5A","#F29E4C","#dc2f02", "#d00000","#370617"))(oo_cols)

ggplot(dat_long_area)+
  geom_boxplot(aes(x = peak_num, y = peak_area, color = HistoMaturity, fill = HistoMaturity))+
  scale_color_manual(values = oo_colors)+
  scale_fill_manual(values = oo_colors)+
  facet_wrap(~ peak_num, scales = "free")

ggplot(dat_long_height)+
  geom_boxplot(aes(x = peak_num, y = peak_height, color = HistoMaturity, fill = HistoMaturity))+
  scale_color_manual(values = oo_colors)+
  scale_fill_manual(values = oo_colors)+
  facet_wrap(~ peak_num, scales = "free")

m1 <- pca(dat[,c(seq(53, 101, by=4))], 7, scale = TRUE, info = "Histostage PCA")
```

### Scree plot

```{r, echo = F}
plotVariance(m1$res$cal, show.labels = TRUE)
```

### Residuals

```{r, echo = F}
plotResiduals(m1, show.labels = FALSE)
```

### PCs 1 and 2

```{r, echo = F, fig.height = 6, fig.width = 6}
#plotScores(m1$res$cal, show.labels = FALSE, cgroup = fp_data_1deriv_1_w$MostProminent)

t <- ggplotly(ggplot(dat, aes(label = barcode))+
  geom_point(aes(m1$res$cal$scores[,1], m1$res$cal$scores[,2], color = dat$HistoMaturity))+
    scale_color_manual(values = oo_colors)+
    labs(title="Scores",
       caption="Source: MACE 2017 Pollock",
       x="Comp 1",
       y="Comp 2",
       color = "Maturity Status")+
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50")+
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50")+
    theme(panel.border = element_rect(fill = NA, color = "grey50"), panel.background = element_blank(), legend.key = element_rect(fill = "white")), width=700, height=500)


htmltools::div(t, align="center")
```


### PCs 1 and 3
```{r, echo = F}
#plotScores(m1$res$cal, c(1,3), show.labels = FALSE, cgroup = fp_data_1deriv_1_w$MostProminent)

t <- ggplotly(ggplot(dat, aes(label = barcode))+
  geom_point(aes(m1$res$cal$scores[,1], m1$res$cal$scores[,3], color = dat$HistoMaturity))+
    scale_color_manual(values = oo_colors)+
  labs(title="Scores",
       caption="Source: MACE 2017 Pollock",
       x="Comp 1",
       y="Comp 3",
       color = "Maturity Status")+
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50")+
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50")+
  theme(panel.border = element_rect(fill = NA, color = "grey50"), panel.background = element_blank(), legend.key = element_rect(fill = "white")), width=700, height=500)


htmltools::div(t, align="center")

```

Let's try a ridge multinomial regression to deal with multicollinearity. All predictors on same scale already, but will mean center. (Dormann et al.)
```{r, warnings = F}
# Prepare data for modeling
## Mean center all variables

peakdat_w <- dat%>% #can I conver this to an operation "accross" for brevity?
  dplyr::transmute(HistoMaturity, GSI.m = GSI - mean(dat$GSI), length.m = length - mean(dat$length), peak1.m = peak_1_height - mean(dat$peak_1_height), peak2.m = peak_2_height - mean(dat$peak_2_height), peak3.m = peak_3_height - mean(dat$peak_3_height), peak4.m = peak_4_height - mean(dat$peak_4_height), peak5.m = peak_5_height - mean(dat$peak_5_height), peak6.m = peak_6_height - mean(dat$peak_6_height), peak7.m = peak_7_height - mean(dat$peak_7_height), peak8.m = peak_8_height - mean(dat$peak_8_height), peak9.m = peak_9_height - mean(dat$peak_9_height), peak10.m = peak_10_height - mean(dat$peak_10_height), peak11.m = peak_11_height - mean(dat$peak_11_height), peak12.m = peak_12_height - mean(dat$peak_12_height), peak13.m = peak_13_height - mean(dat$peak_13_height))

peak_moddf <- subset(peakdat_w, HistoMaturity != "OSNT")
peak_moddf$HistoMaturity <- factor(peak_moddf$HistoMaturity, levels = c("IMM",  "ND",   "DEV",  "VIT",  "PSWN", "SWN",  "PSNT", "SNT"))

peak_moddf <- peak_moddf[,c(1,4:16)]
peak_mat <- model.matrix(HistoMaturity~., peak_moddf)[,-1]
str(peak_mat)
hist_vec <- factor(peak_moddf$HistoMaturity, levels = c("IMM",  "ND",   "DEV",  "VIT",  "PSWN", "SWN",  "PSNT", "SNT"))
# 
# hist_vec <- ifelse(hist_vec == "IMM", 1, ifelse(hist_vec == "ND", 2, ifelse(hist_vec == "DEV", 3, ifelse(hist_vec == "VIT", 4, ifelse(hist_vec == "PSWN", 5, ifelse(hist_vec == "SWN", 6, ifelse(hist_vec == "PSNT", 7, ifelse(hist_vec == "SNT", 8, NA))))))))


#Run a model
## USE glmnet 
#https://cran.r-project.org/web/packages/glmnet/vignettes/glmnet.pdf
#http://www.sthda.com/english/articles/37-model-selection-essentials-in-r/153-penalized-regression-essentials-ridge-lasso-elastic-net/
#https://glmnet.stanford.edu/articles/glmnet.html

#Check out https://www.statlearning.com/

#First, find good lambda value for shrinkage. Defined as the lambda that minimizes the cross-validated prediction error rate.
set.seed(13)
cv <- cv.glmnet(x = peak_mat, y = hist_vec, family = "multinomial") #bootstrap
plot(cv)

```
Now we can find coefficients (or log odds) for optimized lambda (lambda.min). If we exponentiate them, we can find relative risk ratio (odds).
```{r}
coefs <- coef(cv, s = "lambda.min")#Can bootstrap SE for coefs (need to figure out how)
coefs_mat <- rbind(coefs$ND[,1]-coefs$IMM[,1], coefs$DEV[,1]-coefs$IMM[,1], coefs$VIT[,1]-coefs$IMM[,1], coefs$PSWN[,1]-coefs$IMM[,1], coefs$SWN[,1]-coefs$IMM[,1], coefs$PSNT[,1]-coefs$IMM[,1], coefs$SNT[,1]-coefs$IMM[,1]) #extract all coefficients, will need to subtract "ref" coefficient (IMM) from all others.
coefs_mat <- as.data.frame(cbind(coefs_mat, c("ND", "DEV", "VIT", "PSWN", "SWN", "PSNT", "SNT")))

coefs_df <- pivot_longer(coefs_mat, cols = c(2:14), names_to = "peak_num", values_to = "log_odds")
coefs_df <- coefs_df[,-1]
coefs_df$log_odds <- as.numeric(coefs_df$log_odds)
coefs_df$V15 <- as.factor(coefs_df$V15)
coefs_df$peak_num <- as.factor(coefs_df$peak_num)

RRR_df <- coefs_df%>%
  mutate(RRR_m = exp(log_odds))%>%
  rename(maturity = "V15")

```
How to interpret: example, one unit increase in peak.1m intensity is associated with a decrease in log odds of a sample being DEV vs. IMM in the amount of -0.0057. ETC. Also interesting to note a lot of significance in the peak 3, 4, and 6 which we can observe visually by the separation in spectra!


The probability of choosing one outcome category over the probability of the other and is often referred to as relative risk (odds). The relative risk ratios for a unit change in the predictor variable is the right-hand side linear equation exponentiated and therefore equal to exponentiated regression coefficients.
```{r, include = FALSE}

#Try a plot
oo_cols <- length(unique(peak_moddf$HistoMaturity))
oo_colors <- colorRampPalette(c("#370617","#d00000","#dc2f02","#F29E4C","#EFEA5A","#B9E769","#2b9348"))(oo_cols)

plot_1 <- ggplot(subset(RRR_df, peak_num == "peak1.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_2 <- ggplot(subset(RRR_df, peak_num == "peak2.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_3 <- ggplot(subset(RRR_df, peak_num == "peak3.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_4 <- ggplot(subset(RRR_df, peak_num == "peak4.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_5 <- ggplot(subset(RRR_df, peak_num == "peak5.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_6 <- ggplot(subset(RRR_df, peak_num == "peak6.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_7 <- ggplot(subset(RRR_df, peak_num == "peak7.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_8 <- ggplot(subset(RRR_df, peak_num == "peak8.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_9 <- ggplot(subset(RRR_df, peak_num == "peak9.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_10 <- ggplot(subset(RRR_df, peak_num == "peak10.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_11 <- ggplot(subset(RRR_df, peak_num == "peak11.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_12 <- ggplot(subset(RRR_df, peak_num == "peak12.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()

plot_13 <- ggplot(subset(RRR_df, peak_num == "peak13.m"))+
  geom_point(aes(x = RRR_m, y = peak_num, color = maturity), position = position_dodge(width=1), key_glyph = "point")+
  xlim(.90, 1.1)+
  scale_color_manual(values = oo_colors)+
  guides(color = guide_legend(reverse = TRUE, override.aes = list(size=2)))+
  labs(x="Odds (relative risk ratio)",
       y= "Peak number",
       color = "Maturity Status")+
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50")+
  guides(color = "none", fill = "none")+
  theme_classic()
```

```{r, warnings = F, fig.width=6, fig.height=10, }
ggarrange(plot_1, plot_2, plot_3, plot_4, plot_5, plot_6, plot_7, plot_8, plot_9, plot_10, plot_11, plot_12, plot_13, ncol = 2, nrow = 7)

```
For each explantory variable (peak intensity) null and alternative hypotheses can be formuated as:
H0: The relative risk ratio (odds) associated with explantory variable x is equal to 1.
Ha: The relative risk ratio (odds) associated with the explanatory variable x is not equal to 1. 


To examine changes in predicted probability associated with variables, can create small datasets varying on variable while holding other constant. Do this for all peaks in turn while holding others at mean values. First prepare data.
```{r, include = F}

# Peak 1
d_peak1 <- cbind(peak1.m = peak_moddf$peak1.m, peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))

pp.peak1 <- as.data.frame(predict(cv, newx = d_peak1, type = "response", s = "lambda.min"))
pp.peak1 <- cbind(pp.peak1, peak_moddf$peak1.m)

pp.peak1 <- pp.peak1%>%
  transmute(peak = "1", `peak_moddf$peak1.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak1.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#ggplot(pp.peak1)+
  #geom_line(aes(x = peak_diff, y = SWN))

#Peak 2
d_peak2 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = peak_moddf$peak2.m, peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))

pp.peak2 <- as.data.frame(predict(cv, newx = d_peak2, type = "response", s = "lambda.min"))
pp.peak2 <- cbind(pp.peak2, peak_moddf$peak2.m)

pp.peak2 <- pp.peak2%>%
  transmute(peak = "2", `peak_moddf$peak2.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak2.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

# Peak 3
d_peak3 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = peak_moddf$peak3.m, peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))

pp.peak3 <- as.data.frame(predict(cv, newx = d_peak3, type = "response", s = "lambda.min"))
pp.peak3 <- cbind(pp.peak3, peak_moddf$peak3.m)

pp.peak3 <- pp.peak3%>%
  transmute(peak = "3", `peak_moddf$peak3.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak3.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#Peak 4
d_peak4 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = peak_moddf$peak4.m, peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))

pp.peak4 <- as.data.frame(predict(cv, newx = d_peak4, type = "response", s = "lambda.min"))
pp.peak4 <- cbind(pp.peak4, peak_moddf$peak4.m)

pp.peak4 <- pp.peak4%>%
  transmute(peak = "4", `peak_moddf$peak4.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak4.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#Peak 5
d_peak5 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = peak_moddf$peak5.m, peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))

pp.peak5 <- as.data.frame(predict(cv, newx = d_peak5, type = "response", s = "lambda.min"))
pp.peak5 <- cbind(pp.peak5, peak_moddf$peak5.m)

pp.peak5 <- pp.peak5%>%
  transmute(peak = "5", `peak_moddf$peak5.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak5.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#Peak 6
d_peak6 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = peak_moddf$peak6.m, peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))

pp.peak6 <- as.data.frame(predict(cv, newx = d_peak6, type = "response", s = "lambda.min"))
pp.peak6 <- cbind(pp.peak6, peak_moddf$peak6.m)

pp.peak6 <- pp.peak6%>%
  transmute(peak = "6", `peak_moddf$peak6.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak6.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#Peak 7
d_peak7 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = peak_moddf$peak7.m, peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))


pp.peak7 <- as.data.frame(predict(cv, newx = d_peak7, type = "response", s = "lambda.min"))
pp.peak7 <- cbind(pp.peak7, peak_moddf$peak7.m)

pp.peak7 <- pp.peak7%>%
  transmute(peak = "7", `peak_moddf$peak7.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak7.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#Peak 8
d_peak8 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = peak_moddf$peak8.m, peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))


pp.peak8 <- as.data.frame(predict(cv, newx = d_peak8, type = "response", s = "lambda.min"))
pp.peak8 <- cbind(pp.peak8, peak_moddf$peak8.m)

pp.peak8 <- pp.peak8%>%
  transmute(peak = "8", `peak_moddf$peak8.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak8.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#Peak 9
d_peak9 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = peak_moddf$peak9.m, peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))

pp.peak9 <- as.data.frame(predict(cv, newx = d_peak9, type = "response", s = "lambda.min"))
pp.peak9 <- cbind(pp.peak9, peak_moddf$peak9.m)

pp.peak9 <- pp.peak9%>%
  transmute(peak = "9", `peak_moddf$peak9.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak9.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#Peak 10
d_peak10 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = peak_moddf$peak10.m, peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))

pp.peak10 <- as.data.frame(predict(cv, newx = d_peak10, type = "response", s = "lambda.min"))
pp.peak10 <- cbind(pp.peak10, peak_moddf$peak10.m)

pp.peak10 <- pp.peak10%>%
  transmute(peak = "10", `peak_moddf$peak10.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak10.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#Peak 11
d_peak11 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = peak_moddf$peak11.m, peak12.m = mean(peak_moddf$peak12.m), peak13.m = mean(peak_moddf$peak13.m))

pp.peak11 <- as.data.frame(predict(cv, newx = d_peak11, type = "response", s = "lambda.min"))
pp.peak11 <- cbind(pp.peak11, peak_moddf$peak11.m)

pp.peak11 <- pp.peak11%>%
  transmute(peak = "11", `peak_moddf$peak11.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak11.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#Peak 12
d_peak12 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = peak_moddf$peak12.m, peak13.m = mean(peak_moddf$peak13.m))

pp.peak12 <- as.data.frame(predict(cv, newx = d_peak12, type = "response", s = "lambda.min"))
pp.peak12 <- cbind(pp.peak12, peak_moddf$peak12.m)

pp.peak12 <- pp.peak12%>%
  transmute(peak = "12", `peak_moddf$peak12.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak12.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

#Peak 13
d_peak13 <- cbind(peak1.m = mean(peak_moddf$peak1.m), peak2.m = mean(peak_moddf$peak2.m), peak3.m = mean(peak_moddf$peak3.m), peak4.m = mean(peak_moddf$peak4.m), peak5.m = mean(peak_moddf$peak5.m), peak6.m = mean(peak_moddf$peak6.m), peak7.m = mean(peak_moddf$peak7.m), peak8.m = mean(peak_moddf$peak8.m), peak9.m = mean(peak_moddf$peak9.m), peak10.m = mean(peak_moddf$peak10.m), peak11.m = mean(peak_moddf$peak11.m), peak12.m = mean(peak_moddf$peak12.m), peak13.m = peak_moddf$peak13.m)

pp.peak13 <- as.data.frame(predict(cv, newx = d_peak13, type = "response", s = "lambda.min"))
pp.peak13 <- cbind(pp.peak13, peak_moddf$peak13.m)

pp.peak13 <- pp.peak13%>%
  transmute(peak = "13", `peak_moddf$peak13.m`, IMM.1, ND.1, DEV.1, VIT.1, PSWN.1, SWN.1, PSNT.1, SNT.1)%>%
  rename(peak_diff = `peak_moddf$peak13.m`, IMM = IMM.1, ND = ND.1, DEV = DEV.1, VIT = VIT.1, PSWN = PSWN.1, SWN = SWN.1, PSNT = PSNT.1, SNT = SNT.1) #peak diff from mean because centered

# combine probabilities into one data set
pp.all <- rbind(pp.peak1, pp.peak2, pp.peak3, pp.peak4, pp.peak5, pp.peak6, pp.peak7, pp.peak9, pp.peak10, pp.peak11, pp.peak12, pp.peak13)

#melt data to long for ggplot
pp.all_l <- melt(pp.all, id.vars = c("peak", "peak_diff"), value.name = "probability")

```

Now plot these probabilities. This plot shows how the probability of each maturity category changes as the peak intensity increases or decreases away from the average (since peak intensity data were mean centered). This is show individually for each peak. 
```{r, warnings = F}
# Try a plot!!!
oo_colors <- colorRampPalette(c("#2b9348","#B9E769","#EFEA5A","#F29E4C","#dc2f02", "#d00000","#370617"))(oo_cols)

ggplot(pp.all_l)+
  geom_line(aes(x = peak_diff, y = probability, group = variable, color = variable))+
  facet_wrap(~peak, scales = "free_x")+
  scale_color_manual(values = oo_colors)+
  theme_classic()
```


Can use predicted probabilities to help understand the model by converting them back to the scale of the response variable (maturity category). Calculate predicted probabilities for each outcome level using fitted function. We can start by generating the predicted probabilities for the observations in our dataset. Can eventually split data into training and test set.  
```{r, include = F}
pp <- predict(cv, newx = peak_mat, type = "class", s = "lambda.min")

pp_pred <- as.factor(pp)

```

```{r, warnings = F}
# use caret and compute a confusion matrix
mn_conf_mat <-confusionMatrix(data = pp_pred, reference = peak_moddf$HistoMaturity)

mn_conf_mat
```

To look at this in terms of True Positives, True Negatives, False Positives, and False Negatives.
```{r, warnings = F}
#Calculate TP, TN, FP, FN
multi_class_rates <- function(confusion_matrix) {
    true_positives  <- diag(confusion_matrix)
    false_positives <- colSums(confusion_matrix) - true_positives
    false_negatives <- rowSums(confusion_matrix) - true_positives
    true_negatives  <- sum(confusion_matrix) - true_positives -
        false_positives - false_negatives
    return(data.frame(true_positives, false_positives, true_negatives,
                      false_negatives, row.names = names(true_positives)))
}

multi_class_rates(mn_conf_mat$table)
```

Another way to look at this graphically. I will make this prettier...
```{r}
#make dataframe
mn_conf_df <- as.data.frame(mn_conf_mat$table)
test <- mn_conf_df%>%
  group_by(Reference)%>%
  summarize(count = sum(Freq))

mn_conf_df <- left_join(mn_conf_df, test, by = "Reference")

mn_conf_df <- mn_conf_df%>%
  mutate(proportion = Freq/count)

ggplot(mn_conf_df)+
  geom_point(aes(x = Reference, y = Prediction, size = proportion, alpha = proportion == 0.0))+
  scale_alpha_manual(values = c(1,0))+
  scale_size_area()+
  guides(alpha = FALSE)+
  theme_classic()

```

This is really cool actually because we can see what reference categories are most commonly incorrectly classified as and it makes biological sense. For example, PSWN was mostly incorrectly specified as VIT. PSWN = partial-spawning and is biologically very similar to VIT = Vitellogensis. The PSWN stage will have many vitellogenic oocytes present in addition to some hydrated. Higher sample size of VIT in the model which is I think why they are being classified as VIT. 


