---
title: "Raman Spectra Preprocessing"
author: "Esther Goldstein and Morgan Arrington"
date: "1/14/2021"
output: 
  html_document:
  df_print: paged
---

## This document is a WIP for pre-processing methods

Install/open packages
```{r, warning=FALSE,message=FALSE,error=FALSE}
#install.packages('hyperSpec') #package to read in files. Install if you need it
#install.packages("tidyverse") #package for data manipulation, viz and functions. Install if you need it
#install.packages("EMSC") #package for extended multiplicative scatter correction
#install.packages("ggforce") #package for saving figures as pdfs
#install.packages("broom") #for working with tibbles
#install.packages("baseline") #for modified polyfit function, will need to download rtools on CRAN first.
#install.packages("prospectr") #for SG and other functions
library(hyperSpec)
library(tidyverse)
library(EMSC) 
library(ggforce)
library(purrr)
library(broom)
library(baseline)
library(prospectr)
library(plotly)
```



Read in files, format, make into a df, and check with a plot

```{r}
setwd("C:/Users/marri/OneDrive/Documents/AFSC A&G Contract/Raman Spectroscopy/Walleye_Pollock Data/Combined") #will only change wd in this code chunk, have to run together

files<-list.files(path="C:/Users/marri/OneDrive/Documents/AFSC A&G Contract/Raman Spectroscopy/Walleye_Pollock Data/Combined", pattern="WALLEYE_POLLOCK_") #set the file path to the folder with your files.

#read in all the files
spectrumlist <-lapply(files,read.spc)

#make a dataframes out of the portions of the files that I am interested in for analysis
specs<-lapply(spectrumlist, function (x) as.data.frame(cbind(x@wavelength, x@data$spc[1:length(x@data$spc)]))) 

#add file names to the list of files
names(specs)<-lapply(spectrumlist, function (x) x@data$filename) 


#turn list into a dataframe and add a column with the filename
specsdf<-dplyr::bind_rows(specs,.id="id")

colnames(specsdf)<-c("filename","Ramanshift","spc")

p <- ggplot(specsdf)+
  geom_line(aes(x=Ramanshift,y=spc,color=filename))+
  labs(title="Raw raman plot: samples only")+
  theme(legend.position = "none") #This line of code will turn legend off. Too many samples at once and the legend swamps the figure. 

p

```

Subtract airscan
```{r}
#removing airscans from data set for now, still need to figure out if/ how to subtract
#Began taking air scans correctly (2 feet away from all objects) after Bharat's email on 3/2/21. This includes scanning sessions on 3/8/21, 3/10/21, 3/15/21, 4/8/21, and 5/13/21
airscan <- specsdf %>% 
  filter(str_detect(filename, "airscan")) #keep for later

specsdf <- specsdf %>%
  filter(!str_detect(filename, "airscan")) #remove airscan for now

ggplot(airscan)+
  geom_line(aes(x=Ramanshift,y=spc,color=filename))+
  labs(title="Raw raman plot: airscan only")

## Some code below for subtracting airscan if desired
# first filter out airscans that were taken "correctly"
airscan <- airscan %>% 
  filter(filename %in% c("WALLEYE_POLLOCK_airscan_03_08_21", "WALLEYE_POLLOCK_airscan_03_10_21_00000.spc", "WALLEYE_POLLOCK_airscan_03_15_21_00000.spc", "WALLEYE_POLLOCK_airscan_04_08_21_00000.spc", "WALLEYE_POLLOCK_airscan_05_13_21_00000.spc"))

#then take average of these airscans
airscan$filenameA<-airscan$filename #I don't want to lose the entire file name so I made a copy to split

airscan<-splitstackshape::cSplit(airscan,sep="_",splitCols="filenameA")#break the file name up so I can easily group individuals

airscanmeans<-airscan%>%group_by(filenameA_1,filenameA_2,filenameA_3,Ramanshift)%>%summarise(meanspc=mean(spc))


airscanmeans$Sample<-paste(airscanmeans$filenameA_1,airscanmeans$filenameA_2,airscanmeans$filenameA_3,sep="_")

ggplot(airscanmeans,aes(x=Ramanshift,y=meanspc,color=Sample))+
  geom_line()+
  labs(title="Raw raman plot: airscan means")

#Subtract the airscan average from spectra
specsdf_air <- left_join(specsdf, airscanmeans, by = "Ramanshift")%>% #code to subtract airscan from each without writing a loop, restore columns and column names
  mutate(spc = spc-meanspc)%>%
  select(c(filename, Ramanshift, spc))

ggplot(specsdf_air)+
  geom_line(aes(x=Ramanshift,y=spc,color=filename))+
  labs(title="Raw raman plot")+
  theme(legend.position = "none") #This line of code will turn legend off. Too many samples at once and the legend swamps the figure. 

```

Split the data into 2 regions, the fingerprint (200-2100cm-1) and the stretch(2600-3250). 

```{r pressure, echo=FALSE}
#changing fingerprint to 1800 because of spikes
fingerprint<-specsdf_air[specsdf_air$Ramanshift>=200 & specsdf_air$Ramanshift<=1800,] #used to be 200 t0 2100 recommended by Bharat.
stretch<-specsdf_air[specsdf_air$Ramanshift>=2600 & specsdf_air$Ramanshift<=3250,] #2600 to 3250 recommended by Bharat. 
```

On each region, apply EMSE to correct for spectral variation. 
This step is useful if you are trying to compare Raman information without the fluorescence 
background variation.

For the FP region EMSE degree 7 was used. 
```{r}
#needs to be wide format 
fingerprint_w<-fingerprint%>%tidyr::pivot_wider(.,values_from="spc",id_cols="filename",names_from="Ramanshift")

#perform EMSC on the numeric data (i.e. column 1 is the filename)
fingerprint_EMSC<-EMSC(fingerprint_w[,2:ncol(fingerprint_w)], degree=7) # NOTE : MAYBE USE THIS TO ADD TO OVERALL LIST, PLOT EVERYTHING LATER

#combine the corrected values with the filenames so we don't lose info
fingerprint_w_EMSC<-as.data.frame(cbind(fingerprint_w$filename,fingerprint_EMSC$corrected)) 

#change the format to long format for plotting easily in ggplot
fingerprint_EMSC<-fingerprint_w_EMSC%>%tidyr::pivot_longer(.,cols=c(2:ncol(fingerprint_w_EMSC)),names_to="Ramanshift",values_to="spc_EMSC") 

#name the columns
colnames(fingerprint_w_EMSC)[1] <- c("filename")
colnames(fingerprint_EMSC) <- c("filename", "Ramanshift", "spc_EMSC")

#make sure the columns have the right numeric/factor etc assignments
fingerprint_EMSC$filename<-as.factor(fingerprint_EMSC$filename)
fingerprint_EMSC$Ramanshift<-as.numeric(as.character(fingerprint_EMSC$Ramanshift)) 
fingerprint_EMSC$spc_EMSC<-as.numeric(as.character(fingerprint_EMSC$spc_EMSC))

ggplot(fingerprint_EMSC)+
  geom_line(aes(x=Ramanshift,y=spc_EMSC,color=filename))+
  labs(title="Fingerprint region with EMSC")+
  theme(legend.position = "none") #This line of code will turn legend off. Too many samples at once and the legend swamps the figure. 

```

Now stretch region EMSC:

For the stretch r region EMSE degree 5 was used.

Do the same as above but with the stretch section

```{r}
#needs to be wide format 
stretch_w<-stretch%>%tidyr::pivot_wider(.,values_from="spc",id_cols="filename",names_from="Ramanshift")

stretch_EMSC<-EMSC(stretch_w[,2:ncol(stretch_w)], degree=5)

stretch_w_EMSC<-as.data.frame(cbind(stretch_w$filename,stretch_EMSC$corrected))

stretch_EMSC<-stretch_w_EMSC%>%tidyr::pivot_longer(.,cols=c(2:ncol(stretch_w_EMSC)),names_to="Ramanshift",values_to="spc_EMSC")

colnames(stretch_EMSC)<-c("filename","Ramanshift","spc_EMSC")

stretch_EMSC$Ramanshift<-as.numeric(as.character(stretch_EMSC$Ramanshift))
stretch_EMSC$spc_EMSC<-as.numeric(as.character(stretch_EMSC$spc_EMSC))

ggplot(stretch_EMSC)+
  geom_line(aes(x=Ramanshift,y=spc_EMSC,color=filename))+
  labs(title="Stretch region with EMSC")+
  theme(legend.position = "none") #This line of code will turn legend off. Too many samples at once and the legend swamps the figure. 
```

#### Polynomial subtraction used to remove background
Modified polyfit method (Lieber and Mahadevan-Jenson 2003) recommended by Bharat.

Polynomial order 7 was used for FP region and  3 was used for the Stretch region.

```{r}
# How to do this...need a matrix of spectra
filename <- fingerprint_w_EMSC[,1] #save filenames for later
fp_emsc_mat <- as.matrix(fingerprint_w_EMSC)
fp_emsc_mat <- fp_emsc_mat[,-1] #remove first column of filenames
class(fp_emsc_mat) <- "numeric" 

#switch to numeric
# Use modified polyfit function from "baseline" package. More information in Lieber and Mahadevan-Jenson (2003) and Barroso et al. (2018)
polyfit_res <- baseline.modpolyfit(fp_emsc_mat, degree = 7, tol = 0.001, rep = 100)

# Add "baseline" and "corrected" back to dataframe
fp_baseline <- as.data.frame(cbind(filename,polyfit_res$baseline))
fp_corrected <- as.data.frame(cbind(filename,polyfit_res$corrected))

# Make them long format to add back to fingerprint_emsc
fp_w_baseline <- fp_baseline%>%
  tidyr::pivot_longer(.,cols=c(2:ncol(fp_baseline)), values_to="spc_baseline")

fp_w_corrected <- fp_corrected%>%
  tidyr::pivot_longer(.,cols=c(2:ncol(fp_corrected)), values_to="spc_corrected")

# Add back to fingerprint_EMSC
fingerprint_EMSC_mpoly <- cbind(fingerprint_EMSC, as.numeric(fp_w_baseline$spc_baseline), as.numeric(fp_w_corrected$spc_corrected))
colnames(fingerprint_EMSC_mpoly) <- c("filename","Ramanshift", "spc_EMSC", "spc_baseline", "spc_corrected")

# Plot raw spectra, baseline polyfit, and corrected spectra to compare
ggplot(fingerprint_EMSC_mpoly)+
  geom_line(aes(x = Ramanshift, y = spc_EMSC))+
  geom_line(aes(x=fingerprint_EMSC$Ramanshift, y = spc_baseline), color = "blue")+
  geom_line(aes(x = Ramanshift, y = spc_corrected), color = "red")

# Just corrected spectra
ggplot(fingerprint_EMSC_mpoly)+
  geom_line(aes(x = Ramanshift, y = spc_corrected, color = filename)) +
  labs(title="Fingerprint region: EMSC and poly correction")+
  theme(legend.position = "none") #This line of code will turn legend off. Too many samples at once and the legend swamps the figure. 

```

Now Stretch region

Polynomial order 3 
```{r}
# How to do this...need a matrix of spectra
filename <- stretch_w_EMSC[,1] #save filenames for later
stretch_emsc_mat <- as.matrix(stretch_w_EMSC)
stretch_emsc_mat <- stretch_emsc_mat[,-1] #remove first column of filenames
class(stretch_emsc_mat) <- "numeric" 

# Use modified polyfit function from "baseline" package. More information in Lieber and Mahadevan-Jenson (2003) and Barroso et al. (2018)
polyfit_res <- baseline.modpolyfit(stretch_emsc_mat, degree = 3, tol = 0.001, rep = 100)

# Add "baseline" and "corrected" back to dataframe
stretch_w_baseline <- as.data.frame(cbind(filenames,polyfit_res$baseline))
stretch_w_corrected <- as.data.frame(cbind(filenames,polyfit_res$corrected))

# Make them long format to add back to fingerprint_emsc
stretch_baseline <- stretch_w_baseline%>%
  tidyr::pivot_longer(.,cols=c(2:ncol(stretch_w_baseline)), values_to="spc_baseline")

stretch_corrected <- stretch_w_corrected%>%
  tidyr::pivot_longer(.,cols=c(2:ncol(stretch_w_corrected)), values_to="spc_corrected")

# Add back to fingerprint_EMSC
stretch_EMSC_mpoly <- cbind(stretch_EMSC, as.numeric(stretch_baseline$spc_baseline), as.numeric(stretch_corrected$spc_corrected))

colnames(stretch_EMSC_mpoly) <- c("filename","Ramanshift", "spc_EMSC", "spc_baseline", "spc_corrected")

# Plot raw spectra, baseline polyfit, and corrected spectra to compare
ggplot(stretch_EMSC_mpoly)+
  geom_line(aes(x = Ramanshift, y = spc_EMSC))+
  geom_line(aes(x=stretch_EMSC$Ramanshift, y = spc_baseline), color = "blue")+
  geom_line(aes(x = Ramanshift, y = spc_corrected), color = "red")

# Just corrected spectra
ggplot(stretch_EMSC_mpoly)+
  geom_line(aes(x = Ramanshift, y = spc_corrected, color = filename))+
  labs(title="Stretch region: EMSC and poly correction")+
  theme(legend.position = "none") #This line of code will turn legend off. Too many samples at once and the legend swamps the figure. 

```

#### Output .csv files with preprocessed fingerprint and stretch regions just EMSC and mpoly, pre-smoothing
```{r}
# Output file with fingerprint_EMSC_mpoly_sg data
write.csv(x = fingerprint_EMSC_mpoly, file = "fingerprint_EMSC_mpoly.csv")
# Output file with stretch_EMSC_mpoly_sg data
write.csv(x = stretch_EMSC_mpoly, file = "stretch_EMSC_mpoly.csv")

```

#### Apply Savitzky-Golay smoother to reduce noise

Apply Savitzky-Golay smoother to fingerprint region
```{r}
# Savitzky-golay 2nd derivative smoother
fingerprint_EMSC_mpoly_w <-fingerprint_EMSC_mpoly%>%tidyr::pivot_wider(.,values_from="spc_corrected",id_cols="filename",names_from="Ramanshift")

fp_mpoly_mat <- as.matrix(fingerprint_EMSC_mpoly_w)
fp_mpoly_mat <- fp_mpoly_mat[,-1]
class(fp_mpoly_mat) <- "numeric"

spc_sg <- savitzkyGolay(
  X = fp_mpoly_mat,
  m = 0,  #differentiation order
  p = 2,  #polynomial order
  w = 25 #window size, odd
)

# Add "filename" back
fingerprint_EMSC_mpoly_sg_w <- as.data.frame(cbind(filename,spc_sg))

# Make them long format 
fingerprint_EMSC_mpoly_sg <- fingerprint_EMSC_mpoly_sg_w%>%
  tidyr::pivot_longer(.,cols=c(2:ncol(fingerprint_EMSC_mpoly_sg_w)),names_to = "Ramanshift", values_to = "spc_sg")

# Make correct data types
fingerprint_EMSC_mpoly_sg$Ramanshift <- as.numeric(fingerprint_EMSC_mpoly_sg$Ramanshift)
fingerprint_EMSC_mpoly_sg$spc_sg <- as.numeric(fingerprint_EMSC_mpoly_sg$spc_sg)

# Just corrected spectra
ggplot(fingerprint_EMSC_mpoly_sg)+
  geom_line(aes(x = Ramanshift, y = spc_sg, color = filename))+
  labs(title="Fingerprint region: EMSC and poly correction and SG smoothing")+
  theme(legend.position = "none") #This line of code will turn legend off. Too many samples at once and the legend swamps the figure. 

```

Apply Savitzky-Golay smoother to stretch region
```{r}
# Savitzky-golay 2nd derivative smoother
stretch_EMSC_mpoly_w <-stretch_EMSC_mpoly%>%tidyr::pivot_wider(.,values_from="spc_corrected",id_cols="filename",names_from="Ramanshift")

stretch_mpoly_mat <- as.matrix(stretch_EMSC_mpoly_w)
stretch_mpoly_mat <- stretch_mpoly_mat[,-1]
class(stretch_mpoly_mat) <- "numeric"

spc_sg <- savitzkyGolay(
  X = stretch_mpoly_mat,
  m = 0,  #differentiation order
  p = 2,  #polynomial order
  w = 25 #window size, odd
)

# Add "filename" back
stretch_EMSC_mpoly_sg_w <- as.data.frame(cbind(filename,spc_sg))


# Make them long format to add back to fingerprint_emsc
stretch_EMSC_mpoly_sg <- stretch_EMSC_mpoly_sg_w%>%
  tidyr::pivot_longer(.,cols=c(2:ncol(stretch_EMSC_mpoly_sg_w)),names_to = "Ramanshift", values_to = "spc_sg")

# Make correct data types
stretch_EMSC_mpoly_sg$Ramanshift <- as.numeric(stretch_EMSC_mpoly_sg$Ramanshift)
stretch_EMSC_mpoly_sg$spc_sg <- as.numeric(stretch_EMSC_mpoly_sg$spc_sg)

# Just corrected spectra
ggplot(stretch_EMSC_mpoly_sg)+
  geom_line(aes(x = Ramanshift, y = spc_sg, color = filename))+
  labs(title="Stretch region: EMSC and poly correction and SG smoothing")+
  theme(legend.position = "none") #This line of code will turn legend off. Too many samples at once and the legend swamps the figure. 

```

#### Output .csv files with preprocessed fingerprint and stretch regions
```{r}
# Output file with fingerprint_EMSC_mpoly_sg data
write.csv(x = fingerprint_EMSC_mpoly_sg, file = "preprocessed_fingerprint.csv")
write.csv(x = fingerprint_EMSC_mpoly_sg_w, file = "preprocessed_fingerprint_wide.csv")
# Output file with stretch_EMSC_mpoly_sg data
write.csv(x = stretch_EMSC_mpoly_sg, file = "preprocessed_stretch.csv")
write.csv(x = stretch_EMSC_mpoly_sg_w, file = "preprocessed_stretch_wide.csv")

```

#### Apply Savitzky-Golay smoother and 2st derivative for peak height vector

Apply Savitzky-Golay smoother to fingerprint region
```{r}
# Savitzky-golay 2nd derivative smoother
fingerprint_EMSC_mpoly_w <-fingerprint_EMSC_mpoly%>%tidyr::pivot_wider(.,values_from="spc_corrected",id_cols="filename",names_from="Ramanshift")

fp_mpoly_mat <- as.matrix(fingerprint_EMSC_mpoly_w)
fp_mpoly_mat <- fp_mpoly_mat[,-1]
class(fp_mpoly_mat) <- "numeric"

spc_sg <- savitzkyGolay(
  X = fp_mpoly_mat,
  m = 2,  #differentiation order
  p = 2,  #polynomial order
  w = 35 #window size, odd
)

# Add "filename" back
fingerprint_EMSC_mpoly_sg_w <- as.data.frame(cbind(filenames,spc_sg))

# Make them long format to add back to fingerprint_emsc
fingerprint_EMSC_mpoly_sg <- fingerprint_EMSC_mpoly_sg_w%>%
  tidyr::pivot_longer(.,cols=c(2:ncol(fingerprint_EMSC_mpoly_sg_w)),names_to = "Ramanshift", values_to = "spc_sg")

# Make correct data types
fingerprint_EMSC_mpoly_sg$Ramanshift <- as.numeric(fingerprint_EMSC_mpoly_sg$Ramanshift)
fingerprint_EMSC_mpoly_sg$spc_sg <- as.numeric(fingerprint_EMSC_mpoly_sg$spc_sg)

# Just corrected spectra
ggplot(fingerprint_EMSC_mpoly_sg)+
  geom_line(aes(x = Ramanshift, y = spc_sg*-1, color = filenames))+
  labs(title="Fingerprint region: EMSC and poly correction and SG smoothing")+
  theme(legend.position = "none") #This line of code will turn legend off. Too many samples at once and the legend swamps the figure. 

```

Apply Savitzky-Golay smoother to stretch region
```{r}
# Savitzky-golay 2nd derivative smoother
stretch_EMSC_mpoly_w <-stretch_EMSC_mpoly%>%tidyr::pivot_wider(.,values_from="spc_corrected",id_cols="filename",names_from="Ramanshift")

stretch_mpoly_mat <- as.matrix(stretch_EMSC_mpoly_w)
stretch_mpoly_mat <- stretch_mpoly_mat[,-1]
class(stretch_mpoly_mat) <- "numeric"

spc_sg <- savitzkyGolay(
  X = stretch_mpoly_mat,
  m = 2,  #differentiation order
  p = 2,  #polynomial order
  w = 39 #window size, odd
)

# Add "filename" back
stretch_EMSC_mpoly_sg_w <- as.data.frame(cbind(filenames,spc_sg))

# Make them long format to add back to fingerprint_emsc
stretch_EMSC_mpoly_sg <- stretch_EMSC_mpoly_sg_w%>%
  tidyr::pivot_longer(.,cols=c(2:ncol(stretch_EMSC_mpoly_sg_w)),names_to = "Ramanshift", values_to = "spc_sg")

# Make correct data types
stretch_EMSC_mpoly_sg$Ramanshift <- as.numeric(stretch_EMSC_mpoly_sg$Ramanshift)
stretch_EMSC_mpoly_sg$spc_sg <- as.numeric(stretch_EMSC_mpoly_sg$spc_sg)

# Just corrected spectra
ggplot(stretch_EMSC_mpoly_sg)+
  geom_line(aes(x = Ramanshift, y = spc_sg*-1, color = filenames))+
  labs(title="Stretch region: EMSC and poly correction and SG smoothing")+
  theme(legend.position = "none") #This line of code will turn legend off. Too many samples at once and the legend swamps the figure. 

```

#### Output .csv files with preprocessed fingerprint and stretch regions
```{r}
# Output file with fingerprint_EMSC_mpoly_sg data
write.csv(x = fingerprint_EMSC_mpoly_sg, file = "preprocessed_fingerprint_2deriv.csv")
# Output file with stretch_EMSC_mpoly_sg data
write.csv(x = stretch_EMSC_mpoly_sg, file = "preprocessed_stretch_2deriv.csv")

```